Rising from the Ashes II: Can Bar Formation Lead to Abundance Bimodalities?

### 2024-10-09 paper/ms.tex
Update for APOKASC3

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276_alpha.pdf}
  \caption{\textbf{Bimodality in the abundance plane is linked to distinct epochs in simulation.} The upper panels show \MgFe{} as a function of age for our subhalo in TNG. The colors indicate stellar populations at fixed values of \FeH{}, which are the same as in Figure~\ref{fig:fig1}. A gap in the relation occurs at an age of approximately $10.6\,\Gyr$, which we indicate with a vertical dashed line. The effect of the $\alpha$-enhancement is clear, as it separates the stars that form before and after this gap in ages (star particles which formed before $z=1.5$ are $\alpha$-enhanced, which occurs at an age of $\sim9.5\,\Gyr$). The lower panels show on the left the Milky Way and on the center and right the data from TNG but with $15\%$ age errors and $0.01\,\dex$ errors in \MgFe{}. When the simulations are given these errors, we see that the before and after star particles smear such that the two populations significantly overlap in ages. In the $\alpha$-enhanced SoI, two populations emerge in each bin which overlapped in the fiducial distribution. This feature more closely resembles the Milky Way, which displays such populations where the bimodality is strongest -- $\FeH=-0.5$ (blue) and $-0.25$ (orange).}
  \label{fig:alpha}
\end{figure*}

@@

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $15\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC2 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is insignificant, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

@@

\section{Observational Errors}\label{app:obs_err}
In Figure~\ref{fig:alpha}, we assumed observational errors of $15\%$ in age and $0.01\,\dex$ in \MgFe{}. In Figure~\ref{fig:obs_err}, we plot the quoted observational errors of the APOKASC2 (left) and ASPCAP (right) datasets, showing both \FeH{} and \MgFe{} (blue and orange, respectively). We show our $15\%$ age error and $0.01\,\dex$ abundance error assumptions as black lines. For the age error, we used the maximum of the upper and lower estimates from \citet{2018ApJS..239...32P}. Our assumed errors are generally consistent with the errors. Our error in \MgFe{} is a bit smaller than in ASPCAP, but the age estimates are by far the more constraining of the two.

\begin{figure*}
  \centering
  \includegraphics[width=\columnwidth]{fig/obs_error.pdf}
  \caption{The observational errors of the APOKASC2 (left) and ASPCAP dataset (right). We show, on the left, a line indicating a $15\%$ error in observed age and on the right a vertical line indicating a $0.01\,\dex$ error.}
  \label{fig:obs_err}
\end{figure*}

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{fig/392276_alpha.pdf}
  \caption{\textbf{Bimodality in the abundance plane is linked to distinct epochs in simulation.} The upper panels show \MgFe{} as a function of age for our subhalo in TNG. The colors indicate stellar populations at fixed values of \FeH{}, which are the same as in Figure~\ref{fig:fig1}. A gap in the relation occurs at an age of approximately $10.6\,\Gyr$, which we indicate with a vertical dashed line. The effect of the $\alpha$-enhancement is clear, as it separates the stars that form before and after this gap in ages (star particles which formed before $z=1.5$ are $\alpha$-enhanced, which occurs at an age of $\sim9.5\,\Gyr$). The lower panels show on the left the Milky Way and on the center and right the data from TNG but with $12.5\%$ age errors and $0.01\,\dex$ errors in \MgFe{}. When the simulations are given these errors, we see that the before and after star particles smear such that the two populations significantly overlap in ages. In the $\alpha$-enhanced SoI, two populations emerge in each bin which overlapped in the fiducial distribution. This feature more closely resembles the Milky Way, which displays such populations where the bimodality is strongest -- $\FeH=-0.5$ (blue) and $-0.25$ (orange).}
  \label{fig:alpha}
\end{figure*}

@@

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $12.5\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC-3 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is insignificant, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

@@

\section{Observational Errors}\label{app:obs_err}
In Figure~\ref{fig:alpha}, we assumed observational errors of $12.5\%$ in age and $0.01\,\dex$ in \MgFe{}. In Figure~\ref{fig:obs_err}, we plot the quoted observational errors of the APOKASC-3 (left) and ASPCAP (right) datasets, showing both \FeH{} and \MgFe{} (blue and orange, respectively). We show our $12.5\%$ age error and $0.01\,\dex$ abundance error assumptions as black lines. For the age error, we used the maximum of the upper and lower estimates from \citet{2018ApJS..239...32P}. As a dashed line we show our $25\%$ age error cut for stars plotted in Figure~\ref{fig:obs_err}. Our assumed errors are generally consistent with the observed. Our error in \MgFe{} is a bit smaller than in ASPCAP, but the age estimates are by far the more constraining of the two.

\begin{figure*}
  \centering
  \includegraphics[width=\columnwidth]{fig/obs_error.pdf}
  \caption{The observational errors of the APOKASC-3 (left) and ASPCAP dataset (right). We show, on the left, a line indicating a $12.5\%$ error in observed age and on the right a vertical line indicating a $0.01\,\dex$ error. On the left, a dashed line indicates the $25\%$ error cut used for inclusion in Figure~\ref{fig:alpha}.}
  \label{fig:obs_err}
\end{figure*}
